A thesis draft

### 2018-11-30 res/sections/15-ArchitecturalImplementation.tex
Fix English typos

\paragraph*{Low-layer orchestrator}
The low-layer orchestrator job corresponds to the one defined in the ETSI MANO
definition, with the role of managing and orchestrating VNFs and NSs. The
low-layer MANO holds different data repositories where VNF and NS definitions
are stored, various monitoring tools and logs and the network services
repository. The low layer MANO exposes a set of APIs to the upper layer (related
to handling services) MANO, that is able to configure and manage the services.
In other words, the service MANO has the possibility to configure and launch
services to the low-layer orchestrator, providing the necessary information.

\subsection{Low-layer functional components}
In the following section we are going to briefly talk about the role of the most
peculiar components in the defined architecture. A detailed description of the
tasks that the components have to abide in the ETSI MANO architecture is
available in Section~\ref{chap:archimpl:sec:secondattempt:sec:k8s}.
\begin{description}
\item[VNFM] As is possible to see from the
  Figure~\ref{chap:archimpl:sec:fistattempt:img:architecture}, the following
  architecture envisages the possibility of multiple VNFM components, required
  for the control and management of different virtualized functions. In
  particular, the use of different and specialized VNFMs allows to control and
  manage different functional aspects of the control/data plan.
\item[Network service management] in this architectural proposal, it corresponds
  to the brain of the infrastructure MANO, where service instantiation and
  lifecycle management happens. This part interacts with the VNFM/VIM and with
  the RabbitMQ middleware. In this context, the NFVO is involved in chaining the
  required VNFs in order to create the service delivery path. Inside this
  component is possible to identify the Service Lifecycle Management, that
  listens to status report coming from the VNFM and thus generating logs and
  monitoring information.
\item[Monitoring and Information Management] this block gathers data from the
  data layers (identified in the image as the repository), which can be object
  of consultation from other applications.
\end{description}

@@

In our first attempt we started building our infrastructure on top of Ubuntu
16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
would have Openstack registered as PoP, where TOSCA definition would be launched
on Openstack that, instead of deploying hypervisor based virtual machines, it
would have deployed container instances. The VNFs lifecycles would have been
controller through an Element Management System (EMS), transmitting the various
VNF states via the message broker RabbitMQ. Is possible to see the interactions
between elements in a minimum schema available in
Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.

@@

Our first step was to try to use Openstack to create a container orchestration. 
The reason behind this approach was simple: Openbaton, on paper, comes with an 
out-of-the-box support for Openstack. In fact, it natively supports Openstack 
PoPs, giving to us a great possibility to reuse the two frameworks and to ease 
the development phase of our software.
\subsubsection{Devstack: Openstack for developers}
Openstack is composed by modules, giving the possibility to the user who wants 
to install it to choose only the components it really needs. So the first thing 
we decided was to perform a selection of these modules between the 33 available 
in the installation page. The installation of this components span between 
multiple machines (nodes), therefore creating a whole cluster of resources. 
Since we considered too time-consuming to install Openstack in a distributed 
manner, we considered to test first the developer version, called 
\emph{Devstack}. Devstack offers the possibility to install a subset of modules 
in the local machine, and to leverage a Openstack environment able to perform 
the same operations of the production-grade installation. Thus, we starting 
configuring the Devstack installation, installing first of all the Nova 
module, that is required to perform computer operations such as the launch of 
virtual machines, and lately we installed Tacker too.

\paragraph*{Taker}
\begin{figure}[t]
  \centering
  \includegraphics[scale=0.3]{tacker_architecture}
  \caption[Taker architecture schema]{Taker architecture
    schema~\cite{tackerOpenstackwiki}}
\end{figure}
Tacker is a VNFM and a NFVO. It is an official Openstack module, and it has the
ability to deploy, edit, stop and do periodic health checks VNF directly on it.
The developers claim it has the possibility to make deployments to platforms
compatible with Openstack too. It supports VNF definition through the TOSCA
standard, and a proposal for SFC integration has been proposed in
2015~\cite{tackerOpenstackwiki}.

@@

Using Devstack is a simple operation, but the time required to reach a correct 
configuration is very high. In particular, it required from 20 minutes to 1 
hour to install Devstack we a particular configuration, without knowing if the 
final installation was working or not. This trial-and-error approach consumed a 
considerable amount of time, but it was the only possibility we had.

@@

\subsubsection{Openbaton installation}
\paragraph*{Openbaton configuration}
With Devstack ready to be used as NFVI, we started working on Openbaton.
Openbaton currently offers three ways to be deployed:
\begin{itemize}
  \item Normal installation. Openbaton offers official Ubuntu repositories that,
    added into the system, offer a very smooth installation via the
    \verb!aptitude! command line utility, with the only cons of having to wait
    for the automatic Openbaton download, installation and configuration;
  \item Using Vagrant. Vagrant is a software that helps the deployment of
    software, and it offers drivers to deploy virtual virtual machines too;
  \item Docker Compose. At the time of our study, the Docker Compose had three
    different configurations, based on the necessity of the user. A minimal,
    standard and full deployment were offered.
\end{itemize}

Now we are going to do a short description about every installation/deploy method.

\subparagraph*{Installation via repository}
The installation via repositories granted us the possibility to select only the
necessary parts we needed. The back of the medal of this approach was the
impossibility to automate it: in fact, during the installation process user
interaction was required, that we were not able to write down in a scripted
fashion.

\subparagraph*{Deployment via Vagrant}
The deployment via Vagrant confirmed our expectations because it revealed to be
always a smooth operation. The downside of this approach was the lack of
customization and component management. The possibility to remove or add
Openbaton Modules dynamically become later fundamental to us. On top of that,
the Vagrant deployment uses a hypervisor virtualization system (based on Virtual
Box), in contrast with our goal to build a fully container-based system.

\subparagraph*{Deployment via Docker Compose}
The deployment with Docker Compose turned out to be the best of the three
possibilities. The Docker Compose configuration offered an unprecedented
configuration flexibility, with the possibility to deploy only selected modules.
On top of that, this method let us to manage the Openbaton version freely,
having more control on update especially during the development of our
components. Additionally, using this Docker Compose configuration allowed us to
seamlessly integrate the monitoring tools Zabbix, in order to have the
possibility to graphically see the data fetch from Openbaton about the system
state. Finally, a Docker based solution seemed to be the best approach since all
of our components were meant to be put inside a Docker container.

@@

After the initial Openstack and Openbaton configuration we started making the
components talking to each other. First, we tried a simple \verb!iperf! data
transmission with two VNFs deployed by Openbaton as a common hypervisor-based
Openstack instances, as explained in~\cite{openbatonIperf}. With this test
completed, we started tinkering Openstack for using Docker containers instead of
traditional hypervisor instances. When we started developing this solution, we
adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
Unfortunately, we were not able to make it properly work, in particular, it was
not possible to open desired ports on the Docker containers. This lead to the
impossibility to perform packet forwarding between VNFs. On top of that, we
found difficulties regarding the integration of Docker images in TOSCA
definitions and Openstack. These two issues are related: the TOSCA definition
does not have any keyword to specificate in any way that the image used for the
VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
time of the implementation did not performed the image pull from any Docker
registry. The impossibility to easily set the containers ports and the lack of
some support in between the TOSCA definition and the Openstack Zun plugin pushed
us to drop Openstack as a possible solution and to start looking for another
valid VIM component.

@@

First we started playing with Docker Swarm. Docker Swarm is a product directly
made by the Docker Inc., so it seamlessly integrates with other Docker tools
such as Docker Compose. The installation and set-up is pretty straightforward,
and the configuration is almost automatically handed by Docker itself. The
minimum number of nodes in order to create a proper Docker Swarm cluster is of
free nodes, where one must be a manager node (designed to orchestrate the other
nodes in the cluster), while the other can be simple workers. Multiple managers
can coexist in the same cluster, as long as only one effectively operates. The
others act like backup manager in case the one orchestrating goes down for some
reason. The networking is given out-of-the box, and the same holds for
distributed persistence. While this for most of the uses cases can be considered
a pro, in our specific situation the impossibility to tweak the network
configuration was a cons, since it was important for us to have the full control
over the network, especially regarding the SFC implementation, explained better
in Chapter~\ref{chap:vnf_ns_impl}.

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is a orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
the Docker technology. Kubernetes has a architecture based on plug-ins, giving
the possibility to the users to change part of Kubernetes itself, such as the
networking routing strategy or the persistence layer. With this flexibility
Kubernetes offered to us what Docker Swarm was not able to give. Additionally,
we had the possibility to set custom Domain Name System (DNS) lookups and
distribute other internal components in order to avoid requests bottleneck.

@@

Kubernetes does not have a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
We initially deployed from the internal Openstack of the University four nodes
with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines. We
performed the installation using the official Kubernetes Repositories. After the
initial installation, we obtained a configuration similar to the one illustrated
in Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

>>>>

\paragraph*{Low-layer orchestrator}
The low-layer orchestrator job corresponds to the one defined in the ETSI MANO
definition, with the role of managing and orchestrating VNFs and NSs. The
low-layer MANO holds different data repositories where VNF and NS definitions
are stored, alongside with various monitoring tools, logs and network services
repositories. The low layer MANO exposes a set of APIs to the upper layer
(related to handling services) MANO, that is able to configure and manage the
services. In other words, the service MANO has the possibility to configure and
launch services to the low-layer orchestrator, providing the necessary
information.

\subsection{Low-layer functional components}
In the following section we are going to briefly talk about the role of the most
peculiar components in the defined architecture. A detailed description of the
tasks that the components have to abide in the ETSI MANO architecture is
available in Section~\ref{chap:archimpl:sec:secondattempt:sec:k8s}.
\begin{description}
\item[VNFM] As is possible to see from the
  Figure~\ref{chap:archimpl:sec:fistattempt:img:architecture}, the following
  architecture envisages the possibility of multiple VNFM components, required
  for the control and management of different virtualized functions. In
  particular, the use of different and specialized VNFMs allows to control and
  manage different functional aspects of the control/data plan.
\item[Network service management] in this architectural proposal, the Network
  service management corresponds to the brain of the infrastructure MANO, where
  service instantiation and lifecycle management happens. This part interacts
  with the VNFM/VIM and with the RabbitMQ middleware. In this context, the NFVO
  is involved in chaining the required VNFs in order to create the service
  delivery path. Inside this component is possible to identify the Service
  Lifecycle Management, that listens to status report coming from the VNFM and
  it generates logs and monitoring information.
\item[Monitoring and Information Management] this block gathers data from the
  data layers (identified in the image as the repository), which can be object
  of consultation from other applications.
\end{description}

@@

In our first attempt we started building our infrastructure on top of Ubuntu
16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
would have registered Openstack as PoP, where TOSCA definitions would have been
launched on Openstack that, instead of deploying hypervisor based virtual
machines, it would have deployed container instances. The VNFs lifecycles would
have been controlled through an Element Management System (EMS), transmitting
the various VNF states via the message broker RabbitMQ. It is possible to see the
interactions between elements in a minimum schema available in
Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.

@@

Our first step was to try and use Openstack to create a container orchestration.
The reason behind this approach was simple: Openbaton, on the paper, comes with
an out-of-the-box support for Openstack. In fact, it natively supports Openstack
PoPs, giving to us a great possibility to reuse the two frameworks and to ease
the development phase of our software.
\subsubsection{Devstack: Openstack for developers}
Openstack is composed by modules, giving the possibility to the user who wants
to install it to choose only the components it really needs. So the first thing
we decided was to perform a selection of these modules between the 33 available
in the installation page. The installation of this components span between
multiple machines (nodes), therefore creating a whole cluster of resources.
Since we considered too time-consuming to install Openstack in a distributed
manner, we considered to test first the developer version, called
\emph{Devstack}. Devstack offers the possibility to install a subset of modules
in the local machine, and to leverage a Openstack environment able to perform
the same operations of the production-grade installation. Thus, we started
configuring the Devstack installation, installing first of all the Nova module,
that is required to perform computing operations such as the launch of virtual
machines, and lately we installed Tacker too.

\paragraph*{Tacker}
\begin{figure}[t]
  \centering
  \includegraphics[scale=0.3]{tacker_architecture}
  \caption[Taker architecture schema]{Taker architecture
    schema~\cite{tackerOpenstackwiki}}
\end{figure}
Tacker is a VNFM and a NFVO. It is an official Openstack module, and it has the
ability to deploy, edit, stop and do periodic VNF health checks directly on it.
The developers claim it has the possibility to make deployments to platforms
compatible with Openstack too. It supports VNF definition through the TOSCA
standard, and a proposal for SFC integration has been made in
2015~\cite{tackerOpenstackwiki}.

@@

Using Devstack is a simple operation, but the time required to reach a correct 
configuration is very high. In particular, it required from 20 minutes to 1 
hour to install Devstack with a particular configuration, without knowing if the 
final installation was working or not. This trial-and-error approach consumed a 
considerable amount of time, but it was the only possibility we had.

@@

\subsubsection{Openbaton installation}
\paragraph*{Openbaton configuration}
With Devstack ready to be used as NFVI, we started working on Openbaton.
Openbaton currently offers three ways to be deployed:
\begin{itemize}
  \item Normal installation. Openbaton delivers official Ubuntu repositories
    that, added into the system, offer a very smooth installation via the
    \verb!aptitude! command line utility, with the only cons of having to wait
    for the automatic Openbaton download, installation and configuration;
  \item Using Vagrant. Vagrant is a software that helps the software deployment,
    and it offers drivers to deploy virtual virtual machines too;
  \item Docker Compose. At the time of our study, the Docker Compose had three
    different configurations, based on the necessity of the user. A minimal,
    standard and full deployment were offered.
\end{itemize}

A short description about every installation/deploy method will be briefly
illustrated.

\subparagraph*{Installation via repository}
The installation via repository granted us the possibility to select only the
necessary parts we needed. The back of the medal of this approach was the
impossibility to automate it: in fact, during the installation process user
interaction was required, that we were not able to write down in a scripted
fashion.

\subparagraph*{Deployment via Vagrant}
The deployment via Vagrant confirmed our expectations because it revealed to be
always a smooth operation. The downside of this approach was the lack of
customization and component management. The possibility to remove or add
Openbaton modules dynamically became later fundamental to us. On top of that,
the Vagrant deployment uses a hypervisor virtualization system (based on Virtual
Box), in contrast with our goal to build a fully container-based system.

\subparagraph*{Deployment via Docker Compose}
The deployment with Docker Compose turned out to be the best of the three
possibilities. The Docker Compose configuration offered an unprecedented
configuration flexibility, with the possibility to deploy only selected modules.
On top of that, this method let us to manage the Openbaton version freely,
having more control on updates especially during the development of our
components. Additionally, using this Docker Compose configuration allowed us to
seamlessly integrate the monitoring tool Zabbix, in order to have the
possibility to graphically see the data fetch from Openbaton about the system
state. Finally, a Docker based solution seemed to be the best approach since all
of our components were meant to be put inside Docker containers.

@@

After the initial Openstack and Openbaton configuration we started making the
components talking to each other. First, we tried a simple \verb!iperf! data
transmission with two VNFs deployed by Openbaton as a common hypervisor-based
Openstack instances, as explained in~\cite{openbatonIperf}. With this test
completed, we started tinkering Openstack for using Docker containers instead of
traditional hypervisor instances. When we started developing this solution, we
adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
Unfortunately, we were not able to make it properly work, in particular, it was
not possible to open desired ports on the Docker containers. This lead to the
impossibility to perform packet forwarding between VNFs. On top of that, we
found difficulties regarding the integration of Docker images in TOSCA
definitions and Openstack. These two issues are related: the TOSCA definition
does not have any keyword to specificate in any way that the image used for the
VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
time of the implementation did not performed the image pull from any Docker
registry. The impossibility to easily set the containers ports and the lack of
some support between the TOSCA definition and the Openstack Zun plugin pushed us
to drop Openstack as a possible solution and to start looking for another valid
VIM component.

@@

First we started playing with Docker Swarm. Docker Swarm is a product directly
made by the Docker Inc., so it seamlessly integrates with other Docker tools
such as Docker Compose. The installation and set-up is pretty straightforward,
and the configuration is almost automatically handled by Docker itself. The
minimum number of nodes in order to create a proper Docker Swarm cluster is of
three, where one must be a manager node (designed to orchestrate the other
nodes in the cluster), while the others can be simple workers. Multiple managers
can coexist in the same cluster, as long as only one effectively operates. The
others act like backup manager in case the one orchestrating goes down for some
reason. The networking is given out-of-the box, and the same holds for
distributed persistence. While this for most of the uses cases can be considered
a pro, in our specific situation the impossibility to tweak the network
configuration was a cons, since it was important for us to have the full control
over the network, especially regarding the SFC implementation, explained better
in Chapter~\ref{chap:vnf_ns_impl}.

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is an orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
to the users to change part of Kubernetes itself, such as the networking routing
strategy or the persistence layer. With this flexibility Kubernetes offered to
us what Docker Swarm was not able to give. Additionally, we had the possibility
to set custom Domain Name System (DNS) lookups and distribute other internal
components in order to avoid requests bottleneck.

@@

Kubernetes does not have a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
We initially deployed from the internal Openstack of the University four nodes
with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines (that
we later swapped with CentOS 7). We performed the installation using the
official Kubernetes Repositories. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.
